Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets

### 2021-05-18 main2.tex
revise V2:  extended expressiveness

We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
between words and languages, with weighted transducers, 
extending it to infinite alphabets.
% with models of symbolic weighted automata and transducers. 
%
The models introduced in this section generalize 
weighted automata and transducers~\cite{Droste09handbook} 
%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
by labeling each transition with a weight function that takes the 
input and output symbols as parameters, instead of a simple weight value.
These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
but they can return values in an arbitrary semiring, 
where the latter guards are restricted to the Boolean semiring.

@@

Let $\Semiring$ be a commutative and complete semiring, 
 and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be an $\Semiring$-label theory.

\begin{definition}
\label{def:transducer} \label{def:SWT}
A \emph{symbolic-weighted transducer} (\SWT)
over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$ 
is a tuple
$T = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and $\bar{\wei}$ is a 4-uplet of transition functions 
$\wei_{00}$, $\wei_{10}$, $\wei_{01}$, and $\wei_{11}$
%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
from $Q \times Q$ into $\Phi_{\Sigma, \Delta}$.
\end{definition}
%
%A pair of states $\< q, q'>$ such that $\wei_\epsilon(q, q') \neq \zero$
%is called an \emph{$\epsilon$-transition}.

The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
$s$, resp. $t$, being called input word, resp. output word.
%
For convenience, when $\wei_{ij}(q, q') = \eta \in \Phi_{\Sigma, \Delta}$, 
for $q, q' \in Q$ and $i, j \in \{ 0, 1\}$,
we shall sometimes write, overloading the function names:
$\wei_{ij}(q, a, b, q') = \eta(a, b)$ when  $a \in \Sigma$,  $b \in \Delta$.
%
Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
read $a$, resp. $b$, at the current position in the input word, resp. output word;
if $i = 1$, resp. $j = 1$,
increment the current position in the input word (otherwise, do not change it);
change state from $q$ to $q'$.

\noindent
More precisely, the \SWT~$T$ defines a mapping 
from the pairs of words of $\Sigma^* \times \Delta^*$ into~$\Semiring$,
based on the following intermediate function $\weight_T$
defined recursively for every $q, q' \in Q$, 
%$a \in \Sigma$, $b \in \Delta$ 
%for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, 
\[
\begin{array}{lccl}
\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\final(q)}\\
\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
\weight_T(q, au, bv, q') & = & 
   \displaystyle\bigoplus_{q'' \in Q} &
   \wei_{00}(q, a, b, q'') \otimes \weight_T(q'', au, bv, q')\\
  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
    \wei_{10}(q'', a, b, q') \otimes \weight_T(q, u, bv, q'')\\
  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
    \wei_{01}(q'', a, b, q') \otimes \weight_T(q, au, v, q'')\\
  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
    \wei_{11}(q'', a, b, q') \otimes \weight_T(q, u, v, q'')\\
\end{array}
\]

The above sum may be infinite because of the third line.
However, $\weight_T$ is well defined since $\Semiring$ is complete.
%
As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
** SUPPR. $\epsilon$-transition (with $\wei_{00}$) first. **

*** rem. there is no case for $a, \epsilon$ or $\epsilon, b$.
   It can be simulated with $\wei_{10}$ and $\wei_{01}$ on the last letter. ***

%** The application of $\oplus$ can be seen as a ND choice between transitions **
We recall that, by convention, an empty sum with $\oplus$ is $\zero$. 
%
Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
will be ignored in the sum, meaning that there is no possible transition
from state $q$ into state $q'$ while reading $a$ and writing $b$ 
(including the case $a = \epsilon$ and $b = \epsilon$).
%These functions $\phi$ act as guards for the transducer's transitions, 
%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
This is analogous to a transition's guard not satisfied by $\<a, b>$ in 
the case of symbolic transducers.

@@

\begin{example}
Let $\Sigma$ be an alphabet of symbols with timestamps. 
The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.

DTW for sequences of timestamped events **     
\end{example}

\begin{example}
(simpler) pointwise distance between two sequences of timestamped events **
\end{example}

\noindent
The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined similarly as the transducers of Definition~\ref{def:SWT}, 
by simply omitting the output symbols.
%
In this case, the label theory $\bar\Phi$ is reduced to a singleton $\< \Phi_\Sigma>$.
%over $\Sigma$ is reduced to
%a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$ 
is a tuple
$A = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and $\bar{\wei}$ is a pair of transition functions 
$\wei_\epsilon$ and $\wei_\Sigma$ from $Q \times Q$ into 
respectively $\Semiring$ %$\Phi_\epsilon$ 
and $\Phi_\Sigma$.
\end{definition}

\noindent
The above transition functions $\wei_\epsilon$ and $\wei_\Sigma$
can be synthesized as above into a function
$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times Q \to \Semiring$.
%
%When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
%the automaton~$A$ is called \emph{without $\epsilon$-transitions}.

The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***

@@

\begin{proposition}
Given a \SWT $T$ 
over alphabet $\Sigma$, $\Delta$ and a commutative semiring~$\Semiring$, 
and $s \in \Sigma^*$, 
there exists an effectively constructible \SWA 
$A_{T, s}$ over $\Delta$ and $\Semiring$ 
such that for all $t \in \Delta^*$, $A_{T, s}(t) = T(s, t)$.
\end{proposition}
%
\begin{proof}
Let $T = \< Q, \init, \bar{\wei}, \final >$,
let 
$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q 
 \to \Semiring$
be the synthesized form of 
$\bar{\wei} = \< \wei_\epsilon, \wei_\Sigma, \wei_\Delta, \wei_{\Sigma, \Delta}>$,
and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
%$n = |s|$ be the length of $s$.
%
The state set of $A_{T, s}$ will be $Q' = [1..n] \times Q$.
Its state entering weight function is defined by 
$\init'(1, q) = \init(q)$ for all $q \in Q$, 
and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
and its state leaving weight function by 
$\final'(n, q) = \final(q)$ or all $q \in Q$, 
and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.

\noindent
We define transition functions $\wei'_\epsilon$ and $\wei'_\Sigma$ for $A_{T, s}$, 
together with their synthesized form $\wei'$ as follows: 
for all $1 \leq i, j \leq n$, $q, q' \in Q$, %and $a \in \Sigma$, 
and $b \in \Delta$, 
\[
\begin{array}{lclcrcll}
\wei'_\epsilon\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_\epsilon(q, q') & \mathrm{\ie} &
\wei'\bigl(\< i, q>, \epsilon, \< i, q'>\bigr)   & = & \wei(q, \epsilon, \epsilon, q'), & \\ 
%
\wei'_\epsilon\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \phi(s_i) & 
\multicolumn{5}{l}{\mathrm{where~} \phi = \wei_\Sigma(q, q'),}\\
 & & & \mathrm{\ie} &
\wei'\bigl(\< i, q>, \epsilon, \< i+1, q'>\bigr) & = & \wei(q, s_i, \epsilon, q') & 
\mathrm{if~} i < n,\\ 
%
\wei'_\epsilon\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
%\mathrm{\ie} & \wei'(\< i, q>, \epsilon, \< j, q'>)   & = & \zero & & 
\multicolumn{5}{l}{\mathrm{if~} j \neq i, i+1,}\\ 
%
\wei'_\Sigma\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_\Delta(q, q') & \mathrm{\ie} &
\wei'\bigl(\< i, q>, b, \< i, q'>\bigr)          & = & \wei(q, \epsilon, b, q'),\\ 
%
\wei'_\Sigma\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \eta_{s_i} &
\multicolumn{5}{l}{\mathrm{where~} \eta = \wei_{\Sigma, \Delta}(q, q'),}\\
 & & & \mathrm{\ie} & \wei'\bigl(\< i, q>, b, \< i+1, q'>\bigr) & = & \wei(q, s_i, b, q') & 
 \mathrm{if~} i < n,\\ 
%
\wei'_\Sigma\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & \mathrm{\ie} &
\wei'\bigl(\< i, q>, b, \< j, q'>\bigr)     & = & \multicolumn{2}{l}{\zero\ \mathrm{~if~} j \neq i, i+1.}
\end{array}      
\]
We use the assumption that the underlying label theory of $T$ is 
is closed under partial applications (Section~\ref{section:symbols})
in the above construction of $\wei'_\Sigma$.

@@

% gen. to transducers
\begin{proposition}
Given a \SWA $A$ over alphabet $\Sigma$ 
and a commutative and bounded semiring $\Semiring$, 
there exists an effectively constructible \SWA $A'$ 
without $\epsilon$-transitions 
such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
\end{proposition}
%
\begin{proof}  ** Tb revised **   
Let $A = \< Q, \init, \<\wei_\epsilon, \wei_\Sigma>, \final >$.
The automaton $A'$ is $\< Q, \init, \<\wei'_\epsilon, \wei'_\Sigma>, \final >$,
where, for all $q, q' \in Q$,
$\wei'_\epsilon(q, q') = \zero$  
and 
% \[
% \begin{array}{rcccl}
% \wei'_\Sigma(q, q') & = &  & \displaystyle\bigoplus_{q_1 \in Q} &
%                \wei_\Sigma(q, q_1) \otimes  \wei_\epsilon(q_1, q')\\
%                     &   & \oplus  & \displaystyle\bigoplus_{q_2 \in Q} &
%                \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')\\
% \end{array}
% \]
\[
\wei'_\Sigma(q, q') = 
\displaystyle\bigoplus_{q'' \in Q} \wei_\epsilon(q, q'') \otimes \wei_\Sigma(q'', q')
\oplus \displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q'') \otimes \wei_\epsilon(q'', q')
\]
In the above definition of $\wei'_\Sigma$ we use the operator
of product of function of $\Phi_\Sigma$ by $\Semiring$ 
described in Section~\ref{section:symbols}.
%
By definition of $\weight_{A}$ and distributivity of $\otimes$ on $\oplus$, 
** NO. TBC see \cite{Mohri02ijfcs} **
it holds that $\weight_{A}(q, s, q') = \weight_{A'}(q, s, q')$.
\qed
\end{proof}

@@

\subsection{Definition}
Let $\Omega$ be a countable alphabet 
%finite (large) or infinite,
that we assume partitioned into three subsets 
$\Omegai$, $\Omegac$, $\Omegac$,
whose elements are respectively called 
\emph{internal}, \emph{call} and \emph{return} symbols.
% \begin{itemize}
% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
% \end{itemize}
Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative semiring and  
let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
%we shall write respectively 
where $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
for~$\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omegar}$.
%
Moreover, we extend this theory with a set $\Phii$ 
of unary functions in $\Omegai \to \Semiring$,
closed under $\oplus$ and $\otimes$.

@@

\subsection{Best-first Search} 
\label{sec:best}\label{sec:search}
**hypotheses**
Let us assume that the semiring~$\Semiring$ is
commutative, bounded, complete, and total.
and 
** assume label theory is s.t. 
for all $\phi \in \Phii$, 
$\psi \in \Phir$, 
and $\eta \in \Phicr$, 
$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
are computable...

@@

\begin{proposition}
The problem of Symbolic Weighted  parsing 
can be solved in PTIME in the size of the input \SWT, \SWVPA (or \SWTA) and input word, 
and the computation time of the functions of the label theory.
\end{proposition}
%
\begin{proof}
Bar-Hillel construction
\qed  
\end{proof}

\subsection{Application to Automated Music Transcription}
\label{sec:transcription}
Symbolic Automated Music Transcription
and analysis of music performances

\subsubsection{Time Scales}
Real-Time Unit (RTU) = seconds

\noindent 
Musical-Time Unit (MTU) = number of measures

>>>>

We follow the approach of~\cite{Mohri03EDWA} for the computation of distances,
between words and languages, using weighted transducers, 
and extend it to infinite alphabets.
% with models of symbolic weighted automata and transducers. 
%
The models introduced in this section generalize 
weighted automata and transducers~\cite{Droste09handbook} 
%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
by labelling each transition with a weight function (instead of a simple weight value), 
that takes the input and output symbols as parameters. 
These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
but they can return values in an generic semiring, 
whereas the latter guards are restricted to the Boolean semiring.

@@

Let $\Semiring$ be a commutative and complete semiring, 
$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be an $\Semiring$-label theory.

\begin{definition}
\label{def:transducer} \label{def:SWT}
A \emph{symbolic-weighted transducer} (\SWT)
over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$ 
is a tuple
$T = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and $\bar{\wei}$ is a 4-uplet of transition functions 
$\wei_{00}$, $\wei_{10}$, $\wei_{01}$, and $\wei_{11}$
%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
from $Q \times Q$ into $\Phi_{\Sigma, \Delta}$.
\end{definition}

A pair of states $\< q, q'>$ such that $\wei_{00}(q, q') \neq \zero$
is called an \emph{$\epsilon$-transition}.
%
For convenience, when $\wei_{ij}(q, q') = \eta \in \Phi_{\Sigma, \Delta}$, 
for $q, q' \in Q$ and $i, j \in \{ 0, 1\}$,
we shall sometimes write, overloading the function names:
$\wei_{ij}(q, a, b, q') = \eta(a, b)$ when  $a \in \Sigma$,  $b \in \Delta$.

\noindent 
The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
$s$, resp. $t$, being called input word, resp. output word.
%
% \noindent
More precisely, the symbolic-weighted transducer~$T$ defines a mapping 
from the pairs of strings of $\Sigma^* \times \Delta^*$ 
into~$\Semiring$,
based on the following intermediate function $\weight_T$
defined recursively for every $q, q' \in Q$, 
%$a \in \Sigma$, $b \in \Delta$ 
%for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, 
\[
\begin{array}{lccl}
\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
\weight_T(q, au, bv, q') & = & 
   \displaystyle\bigoplus_{q'' \in Q} &
   \wei_{00}(q, a, b, q'') \otimes \weight_T(q'', au, bv, q')\\
  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
    \wei_{10}(q, a, b, q'') \otimes \weight_T(q'', u, bv, q')\\
  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
    \wei_{01}(q, a, b, q'') \otimes \weight_T(q'', au, v, q')\\
  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\\
\end{array}
\]
%
We recall that, by convention, an empty sum with $\bigoplus$ is $\zero$. 
%
The above sum may be infinite because of the $\epsilon$-transitions, % in the first and second lines.
in this case, the infinite sum operators applies since $\Semiring$ is assumed complete.
%As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
%** SUPPR. $\epsilon$-transition (with $\wei_{00}$) first. **
%
%*** rem. there is no case for $a, \epsilon$ or $\epsilon, b$.
%   It can be simulated with $\wei_{10}$ and $\wei_{01}$ on the last letter. ***

%** The application of $\oplus$ can be seen as a ND choice between transitions **
% choose to advance head in input, in output or both (or none)
Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
read $a$, resp. $b$, at the current position in the input word, resp. output word;
if $i = 1$, resp. $j = 1$,
increment the current position in the input word (otherwise, do not change it);
change state from $q$ to $q'$.
The input and output symbols are always read at current positions, 
even during transition that do not increment the head's position.
This is in contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
which ignore the symbol at current position in this case.
In our case, symbols in a infinite alphabet cannot be stored in the finite memory of the transducer, 
whereas considering the current symbol may be useful to compute a transition weight .
%(even when it does not change the head's position, like with $\epsilon$-transitions).
%
The cases $\weight_T(q, au, \epsilon, q')$ and $\weight_T(q, \epsilon, bv, q')$
are missing in the definition of $\weight_T$.
It means that $T$ must avoid configurations where it reached the end of 
the output word and not of the input one, or vice-versa.
This can be done by using $\wei_{10}$ and $\wei_{01}$
before reaching the end of word, and using a special state for this purpose.

Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
will be ignored in the sum, meaning that there is no possible transition
from state $q$ into state $q'$ while reading $a$ and writing $b$ 
(including the case $a = \epsilon$ and $b = \epsilon$).
%These functions $\phi$ act as guards for the transducer's transitions, 
%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
This is analogous to a transition's guard not satisfied by $\<a, b>$ in 
the case of symbolic transducers.

@@

\begin{example}
We can define with a \SWT the computation of a similarity measure 
between timed sequences
similar to dynamic time warping (DTW).

Let $\Semiring$  be the tropical (\emph{min-plus}) semiring of Figure~\ref{fig:semirings} and 
let $\Sigma = \Delta = \mathbb{R}_+$ be sets of timestamps.
We consider a \SWT with one state $q$ and transitions
$\wei_{11}(q, d, d', q) = 
 \wei_{01}(q, d, d', q) = 
 \wei_{10}(q, d, d', q) = |d' - d|$,
and $\wei_{00}(q, d, d', q) = \zero$, for all $d, d' \in \mathbb{R}_+$.
% needs reading input/output symbols even by epsilon-transitions.
The recursive definition of $\weight_T$ correspond to the dynamic programming equations of DTW
for the computation of an optimal match between words, 
the matching cost for two symbols being the 
the time distance between them.
\endex
\end{example}

\begin{example}
% simple pointwise distance between two sequences of timestamped events **
Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$ 
of Figure~\ref{fig:semirings} and 
let $\Sigma = \mathbb{R}_+$ be an input alphabet of event dates
and $\Delta = \{ \mathsf{e}, \mathsf{-} \} \times \mathbb{R}_+$ 
be an output alphabet of symbols with timestamps. 
A symbol $\< \mathsf{e}, d > \in \Delta$ represents an event starting at date $d$, 
and $\< \mathsf{-}, d >$ is a continuation of the previous event.
%More precisely, we let 
%$\Sigma = \{ a, - \} \times \R_+$ and 
This example of $\Delta$ is motivated by the case of music notation, 
where several notated events (notes) can be tied together, 
with a \emph{tie} or a \emph{dot}
(like in $\musQuarter\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\musEighth$ 
 or equivalently~\musQuarterDotted),
meaning that they will be played as a unique sounding event.
%The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.

We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
is to compare a recorded performance $s \in \Sigma^*$
with notated music sheet $t \in \Delta^*$.
One timestamp $d_i \in \Sigma$ may corresponds 
to one notated event $\<\mathsf{e}, d'_i> \in \Sigma$, in which case 
the weight value computed by the \SWT is the time distance between both
(see transitions $\wei_{11}$ below).
%
Moreover, if $\<\mathsf{e}, d'_i>$ is followed by continuations 
$\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
%transitions
\[
\begin{array}{rclcrcl}
\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
\wei_{01}(q_0, d, \< \mathsf{-}, d'>, q_0) & = & \one & &
\wei_{01}(q_1, d, \< \mathsf{-}, d'>, q_0) & = & \one\\
\wei_{10}(q_0, d, b, q_1) & = & \alpha & & \multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
Moreover, it may happen that the performers plays an extra note accidentally, but only once in a row. 
This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
switching from state $q_0$ (normal) to $q_1$ (error).
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances 
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \one$, $\init(q_1) = \zero$, 
$\final(q_0) = \one$, $\final(q_1) = \zero$, 
\endex
\end{example}

\noindent
The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined similarly as the transducers of Definition~\ref{def:SWT}, 
by simply omitting the output symbols.
%
In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
%over $\Sigma$ is reduced to
%a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$ 
is a tuple
$A = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and $\bar{\wei}$ is a pair of transition functions 
$\wei_{0}$, $\wei_{1}$
%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
from $Q \times Q$ into $\Phi_{\Sigma}$.
\end{definition}
A pair of states $\< q, q'>$ such that $\wei_{0}(q, q') \neq \zero$
is called an \emph{$\epsilon$-transition}.

\noindent
As above in the case of \SWT, 
the transition functions $\wei_0$ and $\wei_1$
can be synthesized into a single function
$\wei: Q \times \Sigma \times Q \to \Semiring$.
%
%When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
%the automaton~$A$ is called \emph{without $\epsilon$-transitions}.

%The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***

@@

\begin{proposition}
Given a \SWT $T$ over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$,
and $s \in \Sigma^*$, 
there exists an effectively constructible \SWA 
$A_{T, s}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
such that for all $t \in \Delta^*$, $A_{T, s}(t) = T(s, t)$.
\end{proposition}
%
\begin{proof}
Let $T = \< Q, \init, \bar{\wei}, \final >$,
let 
$\wei: Q \times \Sigma \times \Delta \times Q \to \Semiring$
be the synthesized form of 
$\bar{\wei} = \< \wei_{00}, \wei_{10}, \wei_{01}, \wei_{11}>$,
and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
%$n = |s|$ be the length of $s$.
%
The state set of $A_{T, s}$ will be $Q' = [1..n] \times Q$.
Its state entering weight function is defined by 
$\init'(1, q) = \init(q)$ for all $q \in Q$, 
and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
and its state leaving weight function by 
$\final'(n, q) = \final(q)$ or all $q \in Q$, 
and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.

\noindent
We define transition functions $\wei'_0$ and $\wei'_1$ for $A_{T, s}$, 
together with their synthesized form $\wei'$ as follows: 
for all $1 \leq i, j \leq n$, $q, q' \in Q$, %and $a \in \Sigma$, 
and $b \in \Delta$, 
\[
\begin{array}{lclcr}
\wei'_0\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{00}(q, q')\\ 
%
\wei'_0\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \phi(s_i) & 
\mathrm{if} & i < n, \mathrm{where~} \phi = \wei_{10}(q, q'),\\
%
\wei'_0\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
\mathrm{if} & j \neq i, i+1,\\ 
%
\wei'_1\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{01}(q, q')\\
%
\wei'_1\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \eta_{s_i} &
\mathrm{if} & i < n, \mathrm{where~} \eta = \wei_{11}(q, q'),\\
%
\wei'_1\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
\mathrm{if} & j \neq i, i+1.
\end{array}      
\]
We use the assumption that the underlying label theory of $T$ is 
is closed under partial applications (Section~\ref{section:symbols})
in the above construction of $\wei'_1$.

@@

% gen. to transducers
\begin{proposition}
When $\Semiring$ is commutative, bounded and complete, 
given a \SWA $A$ 
over $\Sigma$, $\Semiring$ and $\bar\Phi$
there exists an effectively constructible \SWA $A'$ 
without $\epsilon$-transitions 
over $\Sigma$, $\Semiring$ and $\bar\Phi$,
such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
\end{proposition}
%
\begin{proof} Let $A = \< Q, \init, \<\wei_0, \wei_1>, \final >$.
The automaton $A'$ is $\< Q, \init, \<\wei'_0, \wei'_1>, \final >$,
where, for all $q, q' \in Q$,
$\wei'_0(q, q') = \zero$  
and 
** Tb revised **   
% \[
% \begin{array}{rcccl}
% \wei'_\Sigma(q, q') & = &  & \displaystyle\bigoplus_{q_1 \in Q} &
%                \wei_\Sigma(q, q_1) \otimes  \wei_\epsilon(q_1, q')\\
%                     &   & \oplus  & \displaystyle\bigoplus_{q_2 \in Q} &
%                \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')\\
% \end{array}
% \]
\[
\wei'_\Sigma(q, q') = 
\displaystyle\bigoplus_{q'' \in Q} \wei_\epsilon(q, q'') \otimes \wei_\Sigma(q'', q')
\oplus \displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q'') \otimes \wei_\epsilon(q'', q')
\]
In the above definition of $\wei'_\Sigma$ we use the operator
of product of function of $\Phi_\Sigma$ by $\Semiring$ 
described in Section~\ref{section:symbols}.
%
By definition of $\weight_{A}$ and distributivity of $\otimes$ on $\oplus$, 
** NO. TBC see \cite{Mohri02ijfcs} **
it holds that $\weight_{A}(q, s, q') = \weight_{A'}(q, s, q')$.
\qed
\end{proof}

@@

\subsection{Definition}
Let $\Omega$ be a countable alphabet 
%finite (large) or infinite,
that we assume partitioned into three subsets 
$\Omegai$, $\Omegac$, $\Omegac$,
whose elements are respectively called 
\emph{internal}, \emph{call} and \emph{return} symbols.
% \begin{itemize}
% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
% \end{itemize}
Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and  
let  $(\Phi_\epsilon, \Phii, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
%we shall write respectively 
where $\Phii$, $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
for~$\Phi_\Omegai$, $\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omegar}$.
%
%Moreover, we extend this theory with a set $\Phii$ 
%of unary functions in $\Omegai \to \Semiring$,
%closed under $\oplus$ and $\otimes$.

@@

\subsection{Best-first Search} 
\label{sec:best}\label{sec:search}
**hypotheses**
Let us assume that the semiring~$\Semiring$ is
commutative, bounded, complete, and total.
and assume an effective label theory.
%s.t. for all $\phi \in \Phii$, 
%$\psi \in \Phir$, 
%and $\eta \in \Phicr$, 
%$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
%$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
%$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
%\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
%are computable...

@@

\begin{proposition}
The problem of Symbolic Weighted  parsing 
can be solved in PTIME in the size of the input \SWT, \SWVPA (or \SWTA) and input word, 
and the computation time of the functions of the label theory.
\end{proposition}
%
%\begin{proof}
%Bar-Hillel construction
%\qed  
%\end{proof}

%\input{transcription}
